Empirical study of land surface effects on local rainfall

### 2017-11-08 chap3_paper_new.Rnw
add files

\begin{keypoints}
\item There are few empirical studies of vegetation feedback on rainfall 
\item Two sites with major step trend landuse change were analysed
\item Statistical change in precipitation detected after bushfire
\end{keypoints}

\begin{abstract}
Empirical evidence of correlations between rainfall and land cover change are scarce due to multiple environmental factors that cannot be controlled. In this study we investigated the relationship using statistical methods on data from best available sources at two sites in Australia. Gridded data of rainfall and tree cover were used as spatially corresponding local conditions. Large scale effects were represented by climatic indicators, such as SOI and IOD. Regression analysis and step trend tests were used to assess the effect of abrupt land surface intervention. At a Queensland site, dramatic tree cover change between 2002 - 2005 did not result in strong statistically significant precipitation changes. On the other hand, results from step trend test on 1998 - 2008 rainfall data in a bushfire affected NSW/VIC region has a better match with the tree cover change map. This indicates the method works better when a abrupt change in the data can be clearly identified. The results from the step trend test implied a positive relationship between the tree cover and the rainfall at 0.1 significance level. 
\end{abstract}

@@

@@

@@

@@

@@

@@

>>>>

\begin{keypoints}
\item There are few empirical studies of vegetation feedback 
\item Two sites with major step trend landuse change were analysed
\item Statistical change in precipitation detected after bushfire
\end{keypoints}

\begin{abstract}
Empirical evidence of correlations between rainfall and land cover change are scarce due to multiple environmental factors that cannot be controlled. In this study we investigated the relationship using statistical methods on data from best available sources at two sites in Australia. Gridded data of rainfall and tree cover were used as spatially corresponding local conditions. Large scale effects were represented by climatic indicators, such as SOI and IOD. Regression analysis and step trend tests were used to assess the effect of abrupt land surface intervention. At a Queensland site, dramatic tree cover change between 2002 - 2005 did not result in statistically significant precipitation change. On the other hand, results from step trend test on 1998 - 2008 rainfall data in a bushfire affected NSW/VIC region has a better match with the tree cover change map. This indicates the method works better when a abrupt change in the data can be clearly identified. The results from the step trend test implied a positive relationship between the tree cover and the rainfall at 0.1 significance level. 
\end{abstract}

\section{Comments to answer by CL FvO and RWV}

\begin{itemize}
  \item comment 2: Do we have to explain why we don't include lag 1 autocorrelation in the model, or use mixed modelling?  
  \item comment 3: should we also report p $<$ 10\%? Now included
  \item comment 5: we seem to be mixing p $<$ 5\% and $\alpha < 0.05$. Changed, Chun can you please check?
\end{itemize}

@@

<<mapping,eval=F, echo = F, results = hide>>=

library(plotKML)
library(raster)
library(rgeos)
library(maptools)
library(fields)
library(RColorBrewer)

## define the boundary of sub-regions
  e1 <- extent(c(1324304.091,1723901.773,-222556.668,80451.265))
  # 400 km by 303 km Qld
  e2 <- extent(c(1346395.141,1534425.527,-1169402.246,-607644.171))
  # 188 km by 561 km NSW
  e11 <- extent(c(145.425, 149.025, -28.475, -25.175))
  e22 <- extent(c(146.675, 149.325, -37.075, -34.675))

  e111 <- extent(c(145.825, 148.625, -28.075, -25.575)) #as(tc_r1,"SpatialPolygons")
  # 306 km by 273.5 km
  e222 <- extent(c(147.075, 148.925, -36.675, -35.075)) #as(tc_r2,"SpatialPolygons")
  # 203 km by 175 km
## the selected regions for mapping

  aus_reg <- readShapePoly("Data\\Geo\\aus_state2011\\simplify_state11_uproj.shp",proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
  aus_reg2 <- gSimplify(aus_reg,0.01,topologyPreserve=T)

  mdb_reg <- readShapePoly("Data\\Geo\\MDB\\mdb.shp",proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
mdb_reg2 <- gSimplify(mdb_reg,0.01,topologyPreserve=TRUE)

  remove(list=c("aus_reg","mdb_reg"))

## Burned area map

fire_0301 <- raster("Data\\Other\\bushfires\\Burnt_Area\\tmpA2003001.burndate.tif")
fire_0301nsw <- crop(fire_0301,e2)
fire_loc <- crop(reproject(fire_0301nsw,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84")),e222)
remove(list=c("fire_0301","fire_0301nsw"))

  pdf("bushfire_nswvic.pdf")
#  jpg("bushfire_nswvic.jpg")
  par(mar=c(3,3,1,6),xpd=T)
  plot(e222,border="white")  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=1.2)
  text(148.4, -36,"Kosciuszko",font=2,cex=0.9)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=1.2)
  text(148, -35.5,"NSW",font=2,cex=1.2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.03,legend.width=1.3,legend.args=list(text="",line=0),lwd.poly=0.1)
legend(149,-36.1,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
 dev.off()

@

<<treeCover,eval=F , echo = F, results = hide>>=

## trend of evi (MODIS product) from GA and ABARES

## tree cover data
  mod44_tc <- brick("Data\\Satellite\\MOD44\\B\\mod44_tc_aus.tif")
  tm <- seq(as.Date("2000-03-06"), as.Date("2010-03-06"), "year")
  mod44_tc <- setZ(mod44_tc, tm, "year")

## tree cover in sub-region
  mod_tc_r11 <- crop(mod44_tc,e1)
  mod_tc_r111 <- reproject(mod_tc_r11,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84"))
  mod_tc_r1 <- crop(mod_tc_r111,e111)

  mod_tc_r22 <- crop(mod44_tc,e2)
  mod_tc_r222 <- reproject(mod_tc_r22,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84"))
 mod_tc_r2 <- crop(mod_tc_r222,e222)

 remove(list=c("mod44_tc","mod_tc_r11","mod_tc_r111","mod_tc_r22","mod_tc_r222"))

## apply t-test to tree cover data
  fun <- function(x,m) {
    var_x <- sd(x)
    mu_1 <- mean(x[1:3])
    if (m==1) {y <- x[6:11]} else {y <- x[4:11]}   
    mu_2 <- mean(y)
    z <- (mu_2-mu_1)/var_x
    if (!is.na(z) & abs(z)>qt(0.1,length(z))) {return(y[1]-x[3])} else {return(NA)}
  }

  mod_tc_r1.z <- calc(mod_tc_r1,function(x) fun(x,1))
  mod_tc_r2.z <- calc(mod_tc_r2,function(x) fun(x,2))

  remove(list=c("mod_tc_r1","mod_tc_r2"))

## create plots
  plotclr <- brewer.pal(6,"PiYG")
#"#C51B7D" "#E9A3C9" "#FDE0EF" "#E6F5D0" "#A1D76A" "#4D9221"

  pdf("tc.pdf", width=10, height= 5)
  par(mar=c(3,3,1,7), mfrow=c(1,2))
  plot(e111)
  image(mod_tc_r1.z,col=plotclr,breaks=c(-70,-50,-20,0,20,50,70),legend=F,add=T)
  plot(e111,add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  image.plot(zlim=c(-3,3),breaks=c(-3,-2,-1,-0.2,0.2,1,2,3),lab.breaks=c(-70,-50,-20,"no tree","0 or",20,50,70),legend.only=T,horizontal=F,col=c("#C51B7D","#E9A3C9","#FDE0EF","white","#E6F5D0","#A1D76A","#4D9221"),legend.shrink=0.6,legend.width=0.7,legend.args=list(text=expression(paste(Delta,"tree (% ground)",sep="")),line=0.5,cex=0.9,font=2),axis.args=list(cex.axis=0.8),graphics.reset=F)
  text(146.0, -25.6, "(a)", font=2)
  plot(e222)
  image(mod_tc_r2.z,col=plotclr,breaks=c(-70,-50,-20,0,20,50,70),legend=F,add=T)
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  text(147.2, -35.1, "(b)", font=2)
  dev.off()

pdf("map.pdf")
par(mar=c(1.5,1.5,1.5,1.5))
plot(mdb_reg2,border="grey",lwd=3)
plot(aus_reg2,border="grey",add=T)
legend(150.2,-36,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
#invisible(text(getSpPPolygonsLabptSlots(aus_reg2), labels=as.character(aus_reg2$NAME), cex=0.7,adj=0.5))
box(lty=1)
plot(e111,col="red",add=T)
plot(e222,col="red",add=T)
text(147.2,-26.8,"QLD",cex=1.5)
text(148,-34.7,"NSW/VIC",cex=1.5)
dev.off()

@

@@

<<climaticIndex,eval=F , echo = F, results = hide>>=

library(reshape)
library(chron)
library(fields)
library(rgdal)
library(raster)
library(maptools)
library(graphics)
library(zoo)
library(RColorBrewer) # creates nice color schemes
# library(classInt)     # finds class intervals for continuous variables

library(mgcv)
library(ncdf4)
library(nnet)

## climatic indicator data

file1 <- "Data\\BoM\\SOI\\SOI.csv"
SOI <- read.csv(file1,sep=",",skip=6)
SOI <- melt(SOI,id="Year")
names(SOI) <- c("Year","Month","SOI")
SOI <- SOI[order(SOI$Year,SOI$Month),]
SOI_ts <- ts(SOI$SOI,start=c(1876,1),end=c(2014,4),frequency=12)
SOI_100a <- window(SOI_ts,start=c(1900,1),end=c(2008,12))
SOI_30a <- window(SOI_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SOI
SOI_100 <- stl(SOI_100a,"period")$time.series[,3]+stl(SOI_100a,"period")$time.series[,2]
SOI_30 <- stl(SOI_30a,"period")$time.series[,3]+stl(SOI_30a,"period")$time.series[,2]

file2 <- "Data\\BoM\\SST\\SST34.cdf"
SST34 <- nc_open(file2)
SST34 <- ncvar_get(SST34,SST34$var[[1]])
SST34_ts <- ts(SST34,start=c(1856,1),end=c(2014,4),frequency=12)
SST34_30a <- window(SST34_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SST34
SST34_30 <- stl(SST34_30a,"period")$time.series[,3]+stl(SST34_30a,"period")$time.series[,2]

file3 <- "Data\\BoM\\SST\\SST3.cdf"
SST3 <- nc_open(file3)
SST3 <- ncvar_get(SST3,SST3$var[[1]])
SST3_ts <- ts(SST3,start=c(1856,1),end=c(2014,4),frequency=12)
SST3_30a <- window(SST3_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SST34
SST3_30 <- stl(SST3_30a,"period")$time.series[,3]+stl(SST3_30a,"period")$time.series[,2]

file4 <- "Data\\BoM\\SST\\SST4.cdf"
SST4 <- nc_open(file4)
SST4 <- ncvar_get(SST4,SST4$var[[1]])
SST4_ts <- ts(SST4,start=c(1856,1),end=c(2014,5),frequency=12)
SST4_30a <- window(SST4_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SST4
SST4_30 <- stl(SST4_30a,"period")$time.series[,3]+stl(SST4_30a,"period")$time.series[,2]

file5 <- "Data\\BoM\\PDO\\PDO.latest_1900-.txt"
pdo_mth <- read.csv(file5,sep="",header=T)
pdo_mth <- melt.data.frame(pdo_mth,id="YEAR")
names(pdo_mth) <- c("Year","Month","PDO")
pdo_mth <- pdo_mth[order(pdo_mth$Year,pdo_mth$Month),]
PDO_ts <- ts(pdo_mth$PDO,start=c(1900,1),end=c(2014,4),frequency=12)
PDO_100a <- window(PDO_ts,start=c(1900,1),end=c(2008,12))
## deseasonalised PDO
PDO_100 <- stl(PDO_100a,"period")$time.series[,3]+stl(PDO_100a,"period")$time.series[,2]

file7 <- "Data\\BoM\\IOD\\iod_jamstec.txt"
IOD <- read.csv(file7,sep="",header=T)
IOD <- IOD[,4]
IOD_ts <- ts(IOD,start=c(1870,1),frequency=12)
IOD_30a <- window(IOD_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised IOD
IOD_30 <- stl(IOD_30a,"period")$time.series[,3]+stl(IOD_30a,"period")$time.series[,2]
# end of read data

@

@@

<<rainData,eval=F , echo = F, results = hide>>=

## 100 years awap data & 30 years
qld_100 <- read.csv("Data\\BOM\\rainfall\\QLD_month_rf.csv",header=T)
qld_30 <- subset(qld_100,as.Date(qld_100[,1],"%d/%m/%Y")>="1979-01-01" & as.Date(qld_100[,1],"%d/%m/%Y")<="2008-12-01")
qld100_mean <- ts(rowMeans(qld_100[,2:2801],na.rm=T,dims=1),start=c(1900,1),frequency=12)
qld30_mean <- window(qld100_mean,start=c(1979,1),end=c(2008,12))

nsw_100 <- read.csv("Data\\BOM\\rainfall\\NSW_month_rf.csv",header=T)
nsw_30 <- subset(nsw_100,as.Date(nsw_100[,1],"%d/%m/%Y")>="1979-01-01" & as.Date(nsw_100[,1],"%d/%m/%Y")<="2008-12-01")
nsw100_mean <- ts(rowMeans(nsw_100[,2:1185],na.rm=T,dims=1),start=c(1900,1),frequency=12)
nsw30_mean <- window(nsw100_mean, start=c(1979,1),end=c(2008,12))

require(MASS)
fitdistr(qld30_mean,"gamma")  ## assume shape=1
fitdistr(nsw30_mean,"gamma")  ## assume shape =8
ks.test(qld30_mean,1,0.03)
ks.test(nsw30_mean,2.4,0.03)

pdf("hist_rain.pdf", width=8, height = 5)
par(mfrow=c(1,2))
 hist(qld30_mean,main="",xlab=expression(paste("rainfall (mm mont",h^-1,")",sep="")))
 text(200,120,"(a)", font=2)
 hist(nsw30_mean,main="",xlab=expression(paste("rainfall (mm mont",h^-1,")",sep="")))
text(200,60,"(b)", font=2)
dev.off()

## autocorrelation in rainfall
pdf("autocorrelation.pdf")
par(mfrow=c(2,2))
  acf(qld30_mean,lag.max=12,main="QLD")
  acf(nsw30_mean,lag.max=12,main="NSW & VIC")

  pacf(qld30_mean,lag.max=12,main="")
  pacf(nsw30_mean,lag.max=12,main="")  
dev.off()

## deseasonalise & detrend rainfall: assume the deseasonalised monthly zonal rainfall is normal distributed and can be used in pearson cross-correlation function
## use stl function in R
  qld30_deseason <- stl(qld30_mean,s.window=length(qld30_mean),t.window=length(qld30_mean))$time.series[,3]
  nsw30_deseason <- stl(nsw30_mean,s.window=length(nsw30_mean),t.window=length(nsw30_mean))$time.series[,3]
  qld100_deseason <- stl(qld100_mean,s.window=length(qld100_mean),t.window=length(qld100_mean))$time.series[,3]
  nsw100_deseason <- stl(nsw100_mean,s.window=length(nsw100_mean),t.window=length(nsw100_mean))$time.series[,3]

pdf("mean_ts.pdf",width=8,height=4)
par(cex.lab=1,cex.axis=1,mfrow=c(1,2))
plot(qld30_deseason,xlab="Year",ylab="monthly rainfall")
abline(v=2003,col="red",lwd=2)
text(2004.6,150,"2003",cex=1.2,col="red")
text(1980.6,200,"(a)",font=2, cex=1.2)
plot(nsw30_deseason,xlab="Year",ylab="monthly rainfall")
abline(v=2003,col="red",lwd=2)
text(2004.6,120,"2003",cex=1.2,col="red")
text(1980.6,150,"(b)", font=2,cex=1.2)
dev.off()

# Figure 5
pdf("cor_qld.pdf", width = 8, height = 8)
par(mfrow=c(2,3),mar=c(5,4,1,1))
ccf(as.vector(SOI_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST34_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO3.4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST4_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100),as.vector(qld100_deseason),lag.max=24,xlim=c(-24,0),xlab="PDO Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100*SOI_100),as.vector(qld100_deseason),lag.max=12,xlim=c(-12,0),xlab="PDO*SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(IOD_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="IOD  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
dev.off()

pdf("cor_nswvic.pdf",width = 8, height = 8)
par(mfrow=c(2,3),mar=c(5,4,1,1))
ccf(as.vector(SOI_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST34_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO3.4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST4_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100),as.vector(nsw100_deseason),lag.max=24,xlim=c(-24,0),xlab="PDO Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100*SOI_100),as.vector(nsw100_deseason),lag.max=12,xlim=c(-12,0),xlab="PDO*SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(IOD_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="IOD  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
dev.off()

@

@@

<<simpleMode,eval=F, echo = F, results = hide>>=

## plotting a hypothesed land cover change model
data1 <- data.frame(year=c(1,2,2,3),tc=c(2,2,1,1))
data2 <- data.frame(year=c(1,2,2,3,4),tc=c(2,2,1,1.5,2))

pdf("tc_simple.pdf", width = 9, height = 4)
par(mar=c(5,6,1,1), mfrow=c(1,2))
plot(data1$year,data1$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",
     axes=F, frame.plot=F)
axis(1,at=c(1,2,3),labels=c("before","LCC","after"),cex.axis=1.3)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=1.3,las=2)
lines(c(1,4),c(2,2),lty=2)
lines(c(1,4),c(1,1),lty=2)
lines(c(2,2),c(0,2.5),lty=2)
text(2.8,2.1,"(a)", font=2)
plot(data2$year,data2$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",axes=F, frame.plot=F)
axis(1,at=c(1,2,4),labels=c("before","LCC","after"),cex.axis=1.3)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=1.3,las=2)
lines(c(1,4),c(2,2),lty=2)
lines(c(1,4),c(1,1),lty=2)
lines(c(2,2),c(0,2.5),lty=2)
text(3.5,2.1,"(b)", font=2)
dev.off()

@

<<fullRegression,eval=F , echo = F, results = hide>>=

## predicting model: land cover is category variable
rain.trend <- function(x,y,m,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))+0.0001

  if (!is.na(mean(var.data$rain))) {
    if (grepl("qld",substitute(y))) {
      var.data <- na.omit(var.data)    
      rain_fit <- gam(rain ~ s(month, k=3)
                  + s(SOI, k=3)
                  + cover  
                  + trd
                  , data=var.data,family=Gamma(link=log))
    } else
     if (grepl("nsw",substitute(y))) {  
      var.data <- na.omit(var.data)    
      rain_fit <- gam(rain ~ s(month, k=3)
                  + s(SOI, k=3)
                  + s(IOD, k=3)
                  + cover  
                  + trd
                  , data=var.data,family=Gamma(link=log))
    }

  return(rain_fit)
  } else {NA}

}

## note: link=log generally has higher r.sq value than link=reverse

## generate variables for the model
trd <- seq(1:360)
month <-rep(1:12,times=30)
mth <- cycle(ts(month,start=c(1979,1),frequency=12))

LC <- c(rep("tree",288),rep("clear",72))
LC_qld4 <- c(rep("tree",312),rep("clear",48)) ## after- : from Jan 2005

var.data <- data.frame(month, trd, SOI = SOI_30a, IOD = IOD_30a, cover = LC)
var.qld4 <- data.frame(month, trd, SOI = SOI_30a, IOD = IOD_30a, cover = LC_qld4)

  coef.rain30.qld4 <- lapply(2:ncol(qld_30), 
                             function(x) rain.trend(x,qld_30,30,
                                                    var.qld4))
  coef.rain30.nswvic <- lapply(2:ncol(nsw_30), 
                               function(x) rain.trend(x,nsw_30,
                                                      30,var.data))

## extract sign of C
  C30_qld <- unlist(lapply(1:length(coef.rain30.qld4),
    function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                       summary(coef.rain30.qld4[[x]])$p.coeff[2],
                       NA)))
  C30_nswvic <- unlist(lapply(1:length(coef.rain30.nswvic),
    function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                       summary(coef.rain30.nswvic[[x]])$p.coeff[2],
                       NA)))

Cp30_qld4 <- unlist(lapply(1:length(coef.rain30.qld4),
   function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                      summary(coef.rain30.qld4[[x]])$p.pv[2],NA)))
Cp30_nswvic <- unlist(lapply(1:length(coef.rain30.nswvic),
    function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                       summary(coef.rain30.nswvic[[x]])$p.pv[2],
                       NA)))

## p value of seasons and climate indices
p_qld <- unlist(lapply(1:length(coef.rain30.qld4),
      function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                         summary(coef.rain30.qld4[[x]])$s.pv[1],
                         NA)))
p_nsw <- unlist(lapply(1:length(coef.rain30.nswvic),
      function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                         summary(coef.rain30.nswvic[[x]])$s.pv[1],
                         NA)))

## extract r.sq of the model
rsq_qld <- unlist(lapply(1:length(coef.rain30.qld4),
     function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                        summary(coef.rain30.qld4[[x]])$r.sq,NA)))
rsq_nsw <- unlist(lapply(1:length(coef.rain30.nswvic),
     function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                        summary(coef.rain30.nswvic[[x]])$r.sq,NA)))

## frame of the two regions (100 years data)
qld_tif <- raster("Data\\BoM\\rainfall\\AWAP\\from1900_FVG\\r19000101_qld.tif")
nsw_tif <- raster("Data\\BoM\\rainfall\\AWAP\\from1900_FVG\\r19000101_nsw.tif")

  dist_Cp30_qld4 <- raster(qld_tif)
  dist_Cp30_qld4[] <- Cp30_qld4
  dist_Cp30_nsw <- raster(nsw_tif)
  dist_Cp30_nsw[] <- Cp30_nswvic
# ## range of predicted values
  predictmax_qld30 <- unlist(lapply(1:length(coef.rain30.qld4),
        function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                          max(coef.rain30.qld4[[x]]$fitted.values),
                           NA)))
  predictmin_qld30 <- unlist(lapply(1:length(coef.rain30.qld4),
        function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",                 min(coef.rain30.qld4[[x]]$fitted.values),NA)))
# 
mean_qld <- lapply(1:length(coef.rain30.qld4), 
          function(x) if(summary(coef.rain30.qld4)[x,2]=="gam") {fitted(coef.rain30.qld4[[x]])
            } else{NA})

## calculate mean residual
  cc <- 1
  qld.sum <- mean_qld[[1]]
  for (i in 2:length(mean_qld)) {
    if(!is.na(mean_qld[[i]])) {
      qld.sum <- qld.sum + mean_qld[[i]]
      cc <- cc+1
    }
  }
  qld.mean <- qld.sum/cc
  qld.mean <- ts(qld.mean,start=c(1979,1),frequency=12)

predictmax_nsw30 <- unlist(lapply(1:length(coef.rain30.nswvic),
        function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                      max(coef.rain30.nswvic[[x]]$fitted.values),
                      NA)))
# 70 - 254.7
  predictmin_nsw30 <- unlist(lapply(1:length(coef.rain30.nswvic),
        function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                      min(coef.rain30.nswvic[[x]]$fitted.values),
                      NA)))
# 18.4 - 62.2

mean_nswvic <- lapply(1:length(coef.rain30.nswvic), 
          function(x) if(summary(coef.rain30.nswvic)[x,2]=="gam") {
              fitted(coef.rain30.nswvic[[x]])} else{NA})

## calculate mean residual
  cc <- 1
  nswvic.sum <- mean_nswvic[[1]]
  for (i in 2:length(mean_nswvic)) {
    if(!is.na(mean_nswvic[[i]])) {
      nswvic.sum <- nswvic.sum + mean_nswvic[[i]]
      cc <- cc+1
    }
  }
  nswvic.mean <- nswvic.sum/cc
  nswvic.mean <- ts(nswvic.mean,start=c(1979,1),frequency=12)

## plots
  pdf("Cp_30yrs.pdf", width=10, height=4.5)
  par(mar=c(3,3,1,6),xpd=T, mfrow=c(1,2))
  plot(e111,xlab="",ylab="")
  image(dist_Cp30_qld4,col=c("red","orange","yellow",
                            "light blue"),
        breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
#   plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  plot(e111,add=T)
text(146, -25.4, "(a)", font=2)
image.plot(zlim=c(0,4),breaks=c(0,1,2,3,4),lab.breaks=c(0,0.05,0.1,0.2,1),legend.only=T,horizontal=F,col=c("red","orange","yellow","light blue"),legend.shrink=0.7,legend.width=0.9,legend.args=list(text="p value",line=0.5,cex=1.2,font=1),axis.args=list(cex.axis=0.8))
legend(148.7,-27.9,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=1)
  plot(e222,xlab="",ylab="") #,border="white")
  image(dist_Cp30_nsw,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  plot(e222,add=T)
  text(147.1, -34.96, "(b)", font=2)
  dev.off()
# 
# 
## plot gam.check
  pdf("qld_check.pdf")
  gam.check(coef.rain30.qld4[[1]])
  dev.off()

  pdf("nswvic_check.pdf")
  gam.check(coef.rain30.nswvic[[which(values(dist_Cp30_nsw)<0.05)[6]]])
  dev.off()

@

@@

%chunk 5: note code has changed
<<chunk5,eval=F , echo = F, results = hide>>=

## without land cover variable for step rend test
rain.trend3 <- function(x,y,m,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))+0.0001

  if (!is.na(mean(var.data$rain))) {
    if (grepl("qld",substitute(y))) {
      rain_fit <- gam(rain ~ trd  # 
                  + s(month,k=3)
                  + s(SOI, k=3)
                  , data=var.data,family=Gamma(link=log))
    } else
    if (grepl("nsw",substitute(y))) {
      rain_fit <- gam(rain ~ trd
                  + s(month,k=3)
                  + s(SOI, k=3)
                  + s(IOD, k=3) 
                  , data=var.data,family=Gamma(link=log))
    }

  return(rain_fit)
  } else {NA}

}
# 
# 
## run model for 30 years
  coef.rain.qld <- lapply(2:ncol(qld_30), 
                function(x) rain.trend3(x,qld_30,30,var.data))
  coef.rain.nswvic <- lapply(2:ncol(nsw_30), 
                function(x) rain.trend3(x,nsw_30,30,var.data))

## explaining power of the partial model
## QLD
  rsq_qld <- unlist(lapply(1:length(coef.rain.qld), 
          function(x) ifelse(summary(coef.rain.qld)[x,2]=="gam",
                             summary(coef.rain.qld[[x]])$r.sq,
                             NA)))
# 
## extract residual
  residual_qld <- lapply(1:length(coef.rain.qld), 
            function(x) if(summary(coef.rain.qld)[x,2]=="gam") {
              coef.rain.qld[[x]]$y-fitted(coef.rain.qld[[x]])
              } else {NA})

# 
# ## NSW/VIC
# 
  rsq_nswvic <- unlist(lapply(1:length(coef.rain.nswvic), 
        function(x) ifelse(summary(coef.rain.nswvic)[x,2]=="gam",
                           summary(coef.rain.nswvic[[x]])$r.sq,
                           NA)))
#

# extract residual
residual_nswvic <- lapply(1:length(coef.rain.nswvic), 
        function(x) if(summary(coef.rain.nswvic)[x,2]=="gam") {
          coef.rain.nswvic[[x]]$y-fitted(coef.rain.nswvic[[x]])
          } else {NA})
saveRDS(residual_nswvic, file="residual_nswvic.RDS")
saveRDS(residual_qld, file="residual_qld.RDS")

## explaining power of the partial model
rain.trend4 <- function(x,y,m,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))+0.0001

  if (!is.na(mean(var.data$rain))) {
    if (grepl("qld",substitute(y))) {
      rain_fit <- gam(rain ~ s(SOI, k=3)
                  , data=var.data,family=Gamma(link=log))
    } else
    if (grepl("nsw",substitute(y))) {
      rain_fit <- gam(rain ~ s(SOI, k=3)
                  + s(IOD, k=3) 
                  , data=var.data,family=Gamma(link=log))
    }

  return(rain_fit)
  } else {NA}

}

## explaining power of index
  coef.soi.qld <- lapply(2:ncol(qld_30), 
                function(x) rain.trend4(x,qld_30,30,var.data))
  coef.soi.nswvic <- lapply(2:ncol(nsw_30), 
                  function(x) rain.trend4(x,nsw_30,30,var.data))
  rsq_soi_qld <- unlist(lapply(1:length(coef.soi.qld), 
          function(x) ifelse(summary(coef.soi.qld)[x,2]=="gam",
                             summary(coef.soi.qld[[x]])$r.sq,NA)))
  rsq_soi_nswvic <- unlist(lapply(1:length(coef.soi.nswvic),
          function(x) ifelse(summary(coef.soi.nswvic)[x,2]=="gam",
                             summary(coef.soi.nswvic[[x]])$r.sq,
                             NA)))

pdf("soi.pdf", width = 10, height = 4)
par(mfrow = c(1,2))
 hist(rsq_soi_qld*100,main="",xlab=expression(paste("% of variability explained (",R^2,")")),cex.lab=1.2)
 text(4.2,400, "(a)", font=2)
 hist(rsq_soi_nswvic*100,main="",xlab=expression(paste("% of variability explained (",R^2,")")),cex.lab=1.2)
text(3.2,300, "(b)", font=2)
dev.off()

@

@@

%chunk 6
<<chunk6,eval=F , echo = F, results = hide>>=

## perform step test on the data
## function to calculate the step trend statistic (Hirsch, 1985)
stepTrend2 <- function(x,y,i){  
    n1 <- 2002-1997
    if (i==1){
      z <- 56  # no. of column
    } else {
    z <- 37 } # close if

  if (x%%z!=0){ #& x%%z!=(z-1)) {
    y.1 <- y[[x]]
    y.2 <- y[[x+1]]
    y.3 <- y[[x+z]]
    y.4 <- y[[x+z+1]]
